---
title: "run_analysis"
author: "Henry Truong"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# "Getting and Cleaning Data" Project

## Task 1: Merging Data from Training Set and Test Set

### Import library

```{r eval=FALSE}
library(tidyr)
library(dplyr)
```

### Download Project Data

**Create file directory for project data**
```{r}
if(!file.exists("Project Data")) {dir.create("Project Data")}
```

**Set current working directory**
```{r eval=FALSE}
myDir = "./Project Data"
setwd(myDir)
```

**Download zip file which is dataset for project**
```{r eval=FALSE}
myUrl = "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
myDest = "./myData.zip"
download.file(myUrl, myDest)
```

**Unzip downloaded zip file**
```{r eval=FALSE}
unzip(myDest)
```

**Set new current working directory**
This step makes codes for importing data somewhat shorter
```{r eval=FALSE}
myDir = "./UCI HAR Dataset"
setwd(myDir)
```

### Import Project Data into RStudio

#### Import Training Dataset

**Read training data on variables X**
```{r eval=FALSE}
myTrainDataX = read.table("./train/X_train.txt")
```

**Read training data on label y**
```{r eval=FALSE}
myTrainDatay = read.table("./train/y_train.txt")
```

**Read training data on volunteer sub**
```{r eval=FALSE}
myTrainSub = read.table("./train/subject_train.txt")
```

#### Import Test Dataset

**Read test data on variables X**
```{r eval=FALSE}
myTestDataX = read.table("./test/X_test.txt")
```

**Read test data on label y**
```{r eval=FALSE}
myTestDatay = read.table("./test/y_test.txt")
```

**Read test data on volunteer sub**
```{r eval=FALSE}
myTestSub = read.table("./test/subject_test.txt")
```

### Make training and test sets understandable

**Add names of variables X to both training and test sets**
```{r eval=FALSE}
names(myTrainDataX) = read.table("./features.txt")[, 2]
names(myTestDataX) = read.table("./features.txt")[, 2]
```

**Add name of label y to both training and test sets**
```{r eval=FALSE}
names(myTrainDatay) = "Activity"
names(myTestDatay) = "Activity"
```

**Add name of volunteer sub to both training and test sets**
```{r eval=FALSE}
names(myTrainSub) = "VolunteerID"
names(myTestSub) = "VolunteerID"
```

### Extract variables whose name contains mean() or std()

**Create boolean vector which determines which variable names contains mean() or std()**
```{r eval=FALSE}
cond = grep("(mean\\(\\)|std\\(\\))", myVarNames)
```

**Extract required subsets from training and test sets**
```{r eval=FALSE}
myCompactTrainDataX = myTrainDataX[, cond]
myCompactTestDataX = myTestDataX[, cond]
```

### Merge all data to get complete table

**Merge all data in training set**
```{r eval=FALSE}
myMergedTrainData = cbind(myTrainSub, myCompactTrainDataX, myTrainDatay)
```

**Merge all data in test set**
```{r eval=FALSE}
myMergedTestData = cbind(myTestSub, myCompactTestDataX, myTestDatay)
```

**Merge training set and test set**
```{r eval=FALSE}
myMergedAllData = rbind(myMergedTrainData, myMergedTestData)
```

### Change labels of Activity from numbers (1, 2, ...) to literal words (WALKING, LAYING, ...)

**Get literal words for Activity**
```{r eval=FALSE}
myActLable = read.table("./activity_labels.txt")[, 2]
```

**Change class of Activity to factor and Set new level**
```{r eval=FALSE}
myMergedAllData$Activity = as.factor(myMergedAllData$Activity)
attributes(myMergedAllData$Activity)$levels = myActLable
```

### Other small changes

**Change class of VolunteerID to factor**
```{r eval=FALSE}
myMergedAllData$VolunteerID = as.factor(myMergedAllData$VolunteerID)
```

**Complete Tast 1 with `myMergedAllData` being result**

## Task 2: Creating table containing mean of variables based on VolunteerID and Activity

### Merge VolunteerID and Activity into 1 column

**Create new table using newly merged column and columns of variable in `myMergedAllData`**
```{r eval=FALSE}
myVolAct = paste(myMergedAllData$VolunteerID, myMergedAllData$Activity, sep = ".")
myNewData = cbind(myVolAct, myMergedAllData[, -c(1, ncol(myMergedAllData))])
```

### Split `myNewData` based on its 1st column and Calculate mean of its variable

**Create new table saving result of calculation applied to each group**
```{r eval=FALSE}
myTable = sapply(split(myNewData[, 2:ncol(myNewData)], myNewData[, 1]), colMeans)
```

### Convert result of calculation into data frame

**Transpose myTable to get 30x66 matrix and convert result to data frame**
```{r eval=FALSE}
myTable = as.data.frame(t(myTable))
```

### Add col VolAct to `myTable` and Remove its row names

**Add col VolAct using row names of `myTable`**
```{r eval=FALSE}
myTBL = cbind(VolAct = row.names(myTable), myTable)
```

**Remove `myTBL`'s row names**
row.names(myTBL) = NULL

### Separate "VolAct" col into "VolunteerID" and "Activity" cols

**Use separate() function**
```{r eval=FALSE}
myTBL = separate(myTBL, VolAct, sep = "\\.", into = c("VolunteerID", "Activity"))
```

**Convert class of VolunteerID to numeric**
```{r eval=FALSE}
myTBL$VolunteerID = as.numeric(myTBL$VolunteerID)
```

**Arrange `myTBL` using VoluteerID column**
```{r eval=FALSE}
myTBL = arrange(myTBL, VolunteerID)
```

### Write result to txt file

```{r eval=FALSE}
write.table(myTBL, file = "../Project Result.txt", sep = ",", row.names = F, col.names = T)
```

**Complete Task 2 with *"Project Result.txt"* being result**